### 2023-02-28 ucIrvine/presentation.tex
Feb 27

\documentclass{beamer}
\usepackage{graphicx}
\title{Uncertainties in precipitation
estimates from space-borne radar observations}
\author{Mircea Grecu}
\institute{NASA Goddard Space Flight Center and Morgan State University}
\date{March 2nd,2023}

\begin{document}
\frame{\titlepage}
\begin{frame}
\frametitle{Sources of uncertainties}
\begin{columns}
%%in the precipitation estimates from space-borne radar observations
\begin{column}{0.45\textwidth}
    Major sources of uncertainty include:
    \begin{itemize}
        \item Variability in the Z-R relationships
        \item Attenuation in the observed radar reflectivity
        \item Variability in the radar beam footprint
        \item Ground clutter
    \end{itemize}
\end{column}
\begin{column}{0.55\textwidth}  %%<--- here
    \begin{figure}
    \begin{center}
     \includegraphics[width=0.9\textwidth]{Figures/fig1.png}
     \end{center}
    \end{figure}
\end{column}
\end{columns}
\end{frame}
\begin{frame}
\frametitle{Variability in the Z-Precipitation Rate relationships}
\begin{itemize}
    \item At least 2 parameters are required to describe
    Particle size distributions (PSDs).
    \item Consequently, the Z-Precipitation Rate relationships require at
    least 2 parameters. The generalized PSD intercept, $N_w=\frac {4^4} {\pi \rho_w} \frac {PWC} {D_m^4}$,
    greatly simplify Z-Rate relationships.
\end{itemize}
\begin{figure}
\begin{center}
\includegraphics[width=0.9\textwidth]{Figures/fig2.png}
\end{center}
\end{figure}
\end{frame}
\begin{frame}
\frametitle{The $N_w$ problem. General considerations}
\begin{columns}
    \begin{column}{0.5\textwidth}
        \begin{itemize}
            \item The $N_w$ parameter greatly simplifies the formulation, but
            not the problem.
            \item It does not solve it though, as $N_w$ still needs to
            estimated independently of the radar observations.
            \item Ground observations may be used to impose constraints on
            $N_w$.
            \item However, such constraints are not very effective for 
            PSD characterized by large mean particle sizes $D_m$.
        \end{itemize}
    \end{column}
    \begin{column}{0.5\textwidth}
        \begin{figure}
            \begin{center}
            \includegraphics[width=0.9\textwidth]{Figures/fig3.png}
            \end{center}
        \end{figure}
        \begin{figure}
            \begin{center}
            \includegraphics[width=0.9\textwidth]{Figures/fig4.png}
            \end{center}
        \end{figure}
    \end{column}
\end{columns}
\end{frame}
\begin{frame}
\frametitle{The $N_w$ problem. Radar profiling considerations}
\begin{columns}
    \begin{column}{0.55\textwidth}
    \begin{itemize}
        \item Additional constraints can be imposed by using radar
        by considering $N_w$ in the radar profiling context.
        \item That is, the attenuation correction process needs to be
        consistent with the $N_w$ parameter.
        $Z(r)=Z_m(r) /PIA$
        $PIA=(1-\epsilon (N_w) q\int_0^r Z_m^\beta(s)ds)^{1/\beta}$
        \item The analytical PIA needs to be consistent with the
        SRT PIA estimate and Ka-band reflectivity observations 
        when available.
    \end{itemize}
    \end{column}
    \begin{column}{0.5\textwidth}
        \begin{figure}
            \begin{center}
            \includegraphics[width=0.9\textwidth]{Figures/fig5.png}
            \end{center}
    \end{figure}
    \end{column}
\end{columns}
\end{frame}
\begin{frame}
\frametitle{The $N_w$ problem. Further considerations}
\begin{itemize}
\item Surface Reference Technique (SRT) PIA is a useful piece of information when reliable.
\item Dual SRT PIA is more reliable than single SRT PIA, except for heavy convection.
\item Over oceans SRT PIA estimates are significantly more reliable than over land.
\item For snow and light rain, SRT PIA estimates are not reliable.
\end{itemize}
\begin{figure}
\begin{center}
\includegraphics[width=0.85\textwidth]{Figures/fig6.png}
\end{center}
\end{figure}
\end{frame}
\begin{frame}
\frametitle{Snow estimation issues}
\begin{columns}
\begin{column}{0.58\textwidth}
    \begin{itemize}
    \item The variability of $N_w$ is still a problem, but there is very little
    attenuation in snow to use the PIA as a constraint.
    \item When dual frequency observations are available, $D_m$ and hence
    $N_w$ may be estimated from the dual frequency reflectivity ratio (DFR).
    \item However, the DFR may be very noisy.
    \item More robust methods, i.e. Machine-Learning based, may be used 
    to derive to incorporate "in-situ" information and provide
    more accurate estimates.
    \end{itemize}
\end{column}
\begin{column}{0.5\textwidth}
\begin{figure}
\begin{center}
\includegraphics[width=0.9\textwidth]{Figures/fig7.png}
\includegraphics[width=0.9\textwidth]{Figures/Fig8.png}
\end{center}
\end{figure}
\end{column}
\end{columns}
\end{frame}    
\end{document}